\textbf{Effective application of machine learning to microbiome based classification problems}

### 2019-10-15 submission/manuscript.Rmd
edited typos and checked word count

For hyperparameter selection, we started with a granular grid search. Then we narrowed and fine-tuned the range of each hyperparameter. A grid search was needed to avoid large variation in prediction performance. For L2-regularized logistic regression, L1 and L2-regularized SVM with linear and radial basis function kernels, we tuned the cost hyperparameter, which controls the regularization strength, where smaller values specify stronger regularization. For SVM with radial basis function kernel, we also tuned the sigma hyperparameter which determines the reach of a single training instance where for a high value of sigma, the SVM decision boundary will be dependent on the points that are closest to the decision boundary. For the decision tree model, we tuned the depth of the tree where the deeper the tree, the more splits it has. For random forest, we tuned the number of features to consider when looking for the best tree split. For XGBoost, we tuned the learning rate and the fraction of samples used for fitting the individual base learners. Performing a grid search for hyperparameter selection might not be feasible for when there are more than two hyperparameters to tune for. In such cases, it is more efficient to use random search or recently developed tools such as Hyperband to identify good hyperparameter configurations [@li_hyperband:_2016].)

@@

**Figure 2. Generalization and classification performance of ML models using AUROC values of all cross validation and testing performances.** The median AUROC for diagnosing individuals with SRN using bacterial abundances was higher than chance (depicted by horizontal line at 0.50) for all the ML models. The predictive performance of random forest model was higher than other ML models, though not significantly (p=0.5). L2-regularized logistic regression, XGBoost, L2-regularized SVM with linear and radial basis function kernel performances were not significantly different from one another. The boxplot shows quartiles at the box ends and the median as the horizontal line in the box. The whiskers show the farthest points that were not outliers. Outliers were defined as those data points that are not within 1.5 times the interquartile ranges. 
\newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

>>>>

For hyperparameter selection, we started with a granular grid search. Then we narrowed and fine-tuned the range of each hyperparameter. A grid search was needed to avoid large variation in prediction performance. For L2-regularized logistic regression, L1 and L2-regularized SVM with linear and radial basis function kernels, we tuned the cost hyperparameter, which controls the regularization strength, where smaller values specify stronger regularization. For SVM with radial basis function kernel, we also tuned the sigma hyperparameter which determines the reach of a single training instance where for a high value of sigma, the SVM decision boundary will be dependent on the points that are closest to the decision boundary. For the decision tree model, we tuned the depth of the tree where the deeper the tree, the more splits it has. For random forest, we tuned the number of features to consider when looking for the best tree split. For XGBoost, we tuned the learning rate and the fraction of samples used for fitting the individual base learners. Performing a grid search for hyperparameter selection might not be feasible for when there are more than two hyperparameters to tune for. In such cases, it is more efficient to use random search or recently developed tools such as Hyperband to identify good hyperparameter configurations [@li_hyperband:_2016].

@@

**Figure 2. Generalization and classification performance of ML models using AUROC values of all cross validation and testing performances.** The median AUROC for diagnosing individuals with SRN using bacterial abundances was higher than chance (depicted by horizontal line at 0.50) for all the ML models. The predictive performance of random forest model was higher than other ML models, though not significantly (p > 0.01). L2-regularized logistic regression, XGBoost, L2-regularized SVM with linear and radial basis function kernel performances were not significantly different from one another. The boxplot shows quartiles at the box ends and the median as the horizontal line in the box. The whiskers show the farthest points that were not outliers. Outliers were defined as those data points that are not within 1.5 times the interquartile ranges. 
\newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

### 2019-10-15 submission/manuscript.tex
edited typos and checked word count

For hyperparameter selection, we started with a granular grid search.
Then we narrowed and fine-tuned the range of each hyperparameter. A grid
search was needed to avoid large variation in prediction performance.
For L2-regularized logistic regression, L1 and L2-regularized SVM with
linear and radial basis function kernels, we tuned the cost
hyperparameter, which controls the regularization strength, where
smaller values specify stronger regularization. For SVM with radial
basis function kernel, we also tuned the sigma hyperparameter which
determines the reach of a single training instance where for a high
value of sigma, the SVM decision boundary will be dependent on the
points that are closest to the decision boundary. For the decision tree
model, we tuned the depth of the tree where the deeper the tree, the
more splits it has. For random forest, we tuned the number of features
to consider when looking for the best tree split. For XGBoost, we tuned
the learning rate and the fraction of samples used for fitting the
individual base learners. Performing a grid search for hyperparameter
selection might not be feasible for when there are more than two
hyperparameters to tune for. In such cases, it is more efficient to use
random search or recently developed tools such as Hyperband to identify
good hyperparameter configurations (47).)

@@

\textbf{Figure 2. Generalization and classification performance of ML
models using AUROC values of all cross validation and testing
performances.} The median AUROC for diagnosing individuals with SRN
using bacterial abundances was higher than chance (depicted by
horizontal line at 0.50) for all the ML models. The predictive
performance of random forest model was higher than other ML models,
though not significantly (p=0.5). L2-regularized logistic regression,
XGBoost, L2-regularized SVM with linear and radial basis function kernel
performances were not significantly different from one another. The
boxplot shows quartiles at the box ends and the median as the horizontal
line in the box. The whiskers show the farthest points that were not
outliers. Outliers were defined as those data points that are not within
1.5 times the interquartile ranges. \newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

>>>>

For hyperparameter selection, we started with a granular grid search.
Then we narrowed and fine-tuned the range of each hyperparameter. A grid
search was needed to avoid large variation in prediction performance.
For L2-regularized logistic regression, L1 and L2-regularized SVM with
linear and radial basis function kernels, we tuned the cost
hyperparameter, which controls the regularization strength, where
smaller values specify stronger regularization. For SVM with radial
basis function kernel, we also tuned the sigma hyperparameter which
determines the reach of a single training instance where for a high
value of sigma, the SVM decision boundary will be dependent on the
points that are closest to the decision boundary. For the decision tree
model, we tuned the depth of the tree where the deeper the tree, the
more splits it has. For random forest, we tuned the number of features
to consider when looking for the best tree split. For XGBoost, we tuned
the learning rate and the fraction of samples used for fitting the
individual base learners. Performing a grid search for hyperparameter
selection might not be feasible for when there are more than two
hyperparameters to tune for. In such cases, it is more efficient to use
random search or recently developed tools such as Hyperband to identify
good hyperparameter configurations (47).

@@

\textbf{Figure 2. Generalization and classification performance of ML
models using AUROC values of all cross validation and testing
performances.} The median AUROC for diagnosing individuals with SRN
using bacterial abundances was higher than chance (depicted by
horizontal line at 0.50) for all the ML models. The predictive
performance of random forest model was higher than other ML models,
though not significantly (p \textgreater{} 0.01). L2-regularized
logistic regression, XGBoost, L2-regularized SVM with linear and radial
basis function kernel performances were not significantly different from
one another. The boxplot shows quartiles at the box ends and the median
as the horizontal line in the box. The whiskers show the farthest points
that were not outliers. Outliers were defined as those data points that
are not within 1.5 times the interquartile ranges. \newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}
